---
title: "Data Pruning"
output: html_document
author: "Franziska M. Willems"
---

# load packages
```{r}
# library(dplyr)
# library(skimr)
# library(data.table)
# library(vtree)
# library(tidyverse)
# library(lme4)

```


```{r}
dat <- read.csv("data/occurrences_full_refined.csv")
```

try fread if read.csv is too slow



```{r}
dat %>%
  group_by(species) %>%
  skim()

dat %>% 
  group_by(species) %>% 
  summarize(N = length(doy),
            mean = mean(doy))
```


# throw out all species with less than 30 datapoints
```{r}
df <- dat
df <- df[df$species != "",]
df <- droplevels(df[ave(df$doy, df$species, FUN = length) > 29, ])

df %>% group_by(species) %>% count() 
```

# thow out all data before 1980
```{r}
hist(df$year)
```

```{r}
df <- droplevels(df[df$year > 1979,])
hist(df$year)
```
# remove doys that are unusually common

```{r}
N_per_doy <- df %>% group_by(doy) %>% summarize(n = n())
ggplot(data = N_per_doy, aes(n, doy, lable = doy))+
  geom_point()+
  geom_text(aes(label=doy),hjust=0, vjust=0)
```



* 1 = 1. January (37589)
* 95 = 5. April (46379)
* 121 = 1. May (75971)
* 163 = 12. June (121456) = Geo Tag der Artenvielfalt

also somewhat more common:
* 181 = 30. June (92207)
* 166 = 15. June (85348)
* 164 = 14. June (86782)

dim(df_sub) = 12544795, 21

```{r}
df %>% filter(doy != 1& 95 & 121 & 163 & 164 & 166 & 181) -> df # this made some trouble, since some doys (e.g. 95) were not removed first
df %>% filter(doy != 95) -> df
df %>% filter(doy != 121) -> df
df %>% filter(doy != 163) -> df 
dim(df)
# Number of dicarded datapoints
12544795 - 13531876  # 281395
```

```{r}
hist(dat$doy, breaks = 365)
hist(df$doy, breaks = 365)

```


# Get summaries and overview of the datset

```{r}
df %>% filter(kingdom == "Animalia") %>% 
  group_by(order, decade) %>% 
  summarize(N = length(doy),
            mean = mean(doy))
```

## check number of datapoints per decade

```{r}
N_per_decade <- df %>% 
  group_by(kingdom, order, species, decade) %>% 
  summarize(N = length(doy),
            mean = mean(doy))
```

```{r}
N_per_decade %>% filter(N > 9) -> N_per_decade_min

N_per_decade_min %>% 
  group_by(species) %>%
 # filter(n() > 3) -> N_per_decade_min # at least 4 decades of data --> leaves 2344 species
  filter(n() > 4) -> N_per_decade_min5 # at least 5 decades of data --> leaves 1272 Species

N_per_decade_min$species <- as.factor(N_per_decade_min$species)
N_per_decade_min$order <- as.factor(N_per_decade_min$order)

# Number of species per order
test <- data.table(N_per_decade_min)
test[ , .(count = length(unique(species))), by = order]
```

## check number of datapoints per decade with 2020 included in the 2010 decade

```{r}
df$decade2 <- df$decade
df$decade2 <- as.factor(df$decade2)
levels(df$decade2)
levels(df$decade2) <- c("1980", "1990", "2000", "2010", "2010")
levels(df$decade2)

N_per_decade2 <- df %>% 
  group_by(kingdom, order, species, decade2) %>% 
  summarize(N = length(doy),
            mean = mean(doy))
```

```{r}
N_per_decade2 %>% filter(N > 9) -> N_per_decade2_min

N_per_decade2_min %>% 
  group_by(species) %>%
  filter(n() > 3) -> N_per_decade2_min # at least 4 decades of data, 2020 included in 2010 


N_per_decade2_min$species <- as.factor(N_per_decade2_min$species)
N_per_decade2_min$order <- as.factor(N_per_decade2_min$order)
```


```{r}
# library(vtree)
vtree(N_per_decade_min[N_per_decade_min$kingdom=="Animalia",], c("order")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
vtree(N_per_decade_min5[N_per_decade_min5$kingdom=="Animalia",], c("order")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)

```

```{r}
ggplot(N_per_decade_min, aes(x = N)) +
  geom_histogram() + 
  facet_grid( ~ kingdom) +
  theme_bw() + xlim(0, 200)

ggplot(N_per_decade_min5, aes(x = N)) +
  geom_histogram() +
  facet_grid( ~ kingdom) +
  theme_bw() +
  xlim(0, 200)

```

# List with species that should be inluded

those that have:
* data from at least 5 decades
* at least 10 datapoints per decade

```{r}
Species_list <- unique(N_per_decade_min5$species)
df %>% filter(species %in% Species_list) -> df_sub
length(levels(as.factor(df_sub$species)))
df_sub$species <- as.factor(df_sub$species)
```

# the same when 2020 is included in the 2010 decade
```{r}
Species_list2 <- unique(N_per_decade2_min$species)

memory.limit(size=999999999)
df %>% filter(species %in% Species_list2) -> df_sub2

saveRDS(df_sub2, file = "df_sub2.RDS")
length(levels(as.factor(df_sub2$species))) # 2046
df_sub2$species <- as.factor(df_sub2$species)
```

## number of datapoints per animal order and species

```{r}
vtree(df_sub[df_sub$kingdom=="Animalia",], c("order")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
vtree(df_sub[df_sub$order=="Coleoptera",], c("species")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
vtree(df_sub[df_sub$order=="Diptera",], c("species")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
vtree(df_sub[df_sub$order=="Hymenoptera",], c("species")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
vtree(df_sub[df_sub$order=="Lepidoptera",], c("species")) #, fillcolor = c( order = "#e7d4e8")) #, horiz = FALSE)
length(levels(as.factor(df_sub$species)))
N_per_species <- df_sub %>% group_by(species) %>% count() 
```
# Overview plots of data

## map with data points

```{r}
theme_set(theme_bw())
# library(rworldmap)
# library(rgeos)
# library(sf)
world <- getMap(resolution = "high")
world <- st_as_sf(world)
```



```{r}
ggplot(data = world)+
  geom_sf()+
  geom_point(data = df_sub, aes(y = decimalLatitude, x = decimalLongitude))+
  coord_sf(xlim = c(0,16), ylim = c(45,55))
```


```{r}
ggplot(data = df_sub, aes(x = year, y = doy, colour = kingdom)) +
         geom_point()+
         geom_smooth(method = "lm")
```
```{r}
ggplot(data = df_sub, aes(x = year, y = doy, colour = order)) +
#         geom_point()+
         geom_smooth(method = "lm")
```

```{r}
ggplot(data = df_sub[df_sub$kingdom=="Animalia",], aes(x = year, y = doy, group = species)) +
         geom_point(alpha = 0.5)+
         geom_smooth(method = "lm")+
  facet_wrap(~order)
```
```{r}
ggplot(data = df_sub[df_sub$order=="Coleoptera",], aes(x = year, y = doy)) +
         geom_point(alpha = 0.5)+
         geom_smooth(method = "lm")+
  facet_wrap(~species)
```
```{r}
ggplot(data = df_sub[df_sub$order=="Diptera",], aes(x = year, y = doy)) +
         geom_point(alpha = 0.5)+
         geom_smooth(method = "lm")+
  facet_wrap(~species)
```


```{r}
ggplot(data = df_sub[df_sub$order=="Hymenoptera",], aes(x = year, y = doy)) +
         geom_point(alpha = 0.5)+
         geom_smooth(method = "lm")+
         geom_smooth(data = df_sub[df_sub$order=="Hymenoptera"&df_sub$year > 2009,], method = "lm", col = "red")+
  facet_wrap(~species)
```



```{r}
ggplot(data = df_sub[df_sub$order=="Lepidoptera",], aes(x = year, y = doy)) +
         geom_point(alpha = 0.5)+
         geom_smooth(method = "lm")+
         geom_smooth(data = df_sub[df_sub$order=="Lepidoptera"&df_sub$year > 2009,], method = "lm", col = "red")+
  facet_wrap(~species)
```




