---
title: "Get CRU data as RasterBrick Object"
output: html_notebook
---
for instructions see: 
https://www.benjaminbell.co.uk/2018/01/getting-climate-data.html

```{r}
# Load packages
library(raster)
library(ncdf4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(data.table)
```
# load Data
```{r}
df_sub <- readRDS(file = "/Franziska/CleanWorkBench/GBIF_Data/df_sub2.RDS")
```


## round latitude/longitude to .00, .25, .50, .75
```{r}
# df_sub$lat <- round(df_sub$decimalLatitude, 2)
# df_sub$long <- round(df_sub$decimalLongitude, 2)

# library(plyr)
# df_sub$lat <-round_any(df_sub$lat,0.25)
# df_sub$long <-round_any(df_sub$long,0.25)
```




```{r}
# Load the CRU TS temperature dataset into R 

temp1971_1980 <- brick("download/cru_ts4.05.1971.1980.tmp.dat.nc", varname="tmp")
temp1981_1990 <- brick("download/cru_ts4.05.1981.1990.tmp.dat.nc", varname="tmp")
temp1991_2000 <- brick("download/cru_ts4.05.1991.2000.tmp.dat.nc", varname="tmp")
temp2001_2010 <- brick("download/cru_ts4.05.2001.2010.tmp.dat.nc", varname="tmp")
temp2011_2020 <- brick("download/cru_ts4.05.2011.2020.tmp.dat.nc", varname="tmp")

#nlayers(temp1981_1990)
names(temp1981_1990)

```
The above code loads the NetCDF file as a "RasterBrick" object (similar to an array) into memory using the raster package

The temperature/precipitation data files has have variables "tmp" ("pre") and "stn", 
"pre" = total precipitation (mm/month), and 
"stn" = the number of climate stations that were used for the grid square.

dimensions = refesr to longitude and latitude, so  the file contains 
climate data for 720 x 360 grid squares covering the entire globe. 
crs = the format of the coordinates system. 
The third dimension represents time, covering each month between January 1901 and December 2016. 
This represents 1428 data entries, which can be confirmed by multiplying the number of years (119) 
by the number of months (12), which equals 1428 

It is possible to plot the information directly from a "RasterBrick" object. 
E.g. a map showing global precipitation for January 1901 just to see how the data looks: 


# Look at Germany area only

crop roughly
- longitude 6 - 15 
- latitude 47 - 55

```{r}
GER.area <- extent(6, 15, 47, 55)
temp1971_1980 <- crop(temp1971_1980, GER.area)
temp1981_1990 <- crop(temp1981_1990, GER.area)
temp1991_2000 <- crop(temp1991_2000, GER.area)
temp2001_2010 <- crop(temp2001_2010, GER.area)
temp2011_2020 <- crop(temp2011_2020, GER.area)

```



# remove NAs in dataset and create df with only "sites", and decaded dfs
```{r}
D <- df_sub %>% filter(!is.na(decimalLatitude)) # %>% select(kingdom, order, species, id.grp, decimalLatitude, decimalLongitude, decade, year, doy)

D_1971_1980 <- D %>% filter(year > 1970 & year <= 1980)
D_1981_1990 <- D %>% filter(year > 1980 & year <= 1990)
D_1991_2000 <- D %>% filter(year > 1990 & year <= 2000)
D_2001_2010 <- D %>% filter(year > 2000 & year <= 2010)
D_2011_2020 <- D %>% filter(year > 2010 & year <= 2020)
```

# Extract data for 1970s (that is ONLY 1980 in this case)
## Extract Temp values from RasterBrick
```{r}
data_to_match = D_1971_1980
temp_raster = temp1971_1980

coordinates(data_to_match)= ~ decimalLongitude + decimalLatitude
rasValue = raster::extract(temp_raster, data_to_match) # layer = c(1:12)
combinePointValue = cbind(data_to_match,rasValue)

DT_1970s <- as.data.frame(combinePointValue)
# colnames(DT_1980s) <- gsub("^c\\..*", "Temp_1981_3", colnames(DT_1980s))
```

## aggregate monthly data
```{r}
DT_1970s %>% mutate(Temp1971 = rowMeans(select(DT_1970s, starts_with("X1971.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1972 = rowMeans(select(DT_1970s, starts_with("X1972.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1973 = rowMeans(select(DT_1970s, starts_with("X1973.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1974 = rowMeans(select(DT_1970s, starts_with("X1974.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1975 = rowMeans(select(DT_1970s, starts_with("X1975.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1976 = rowMeans(select(DT_1970s, starts_with("X1976.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1977 = rowMeans(select(DT_1970s, starts_with("X1977.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1978 = rowMeans(select(DT_1970s, starts_with("X1978.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1979 = rowMeans(select(DT_1970s, starts_with("X1979.")))) -> DT_1970s
DT_1970s %>% mutate(Temp1980 = rowMeans(select(DT_1970s, starts_with("X1980.")))) -> DT_1970s

colnames(DT_1970s)
DT_1970s %>% select(-starts_with("X")) -> DT_1970s
colnames(DT_1970s)

DT_1970s %>%
  mutate(Temp = case_when(
    year == 1971 ~ DT_1970s$Temp1971, 
    year == 1972 ~ DT_1970s$Temp1972, 
    year == 1973 ~ DT_1970s$Temp1973, 
    year == 1974 ~ DT_1970s$Temp1974, 
    year == 1975 ~ DT_1970s$Temp1975, 
    year == 1976 ~ DT_1970s$Temp1976, 
    year == 1977 ~ DT_1970s$Temp1977, 
    year == 1978 ~ DT_1970s$Temp1978, 
    year == 1979 ~ DT_1970s$Temp1979, 
    year == 1980 ~ DT_1970s$Temp1980)) -> DT_1970s                    

```

#############


# Extract data for 1980s
## Extract Temp values from RasterBrick
```{r}
data_to_match = D_1981_1990
temp_raster = temp1981_1990

coordinates(data_to_match)= ~ decimalLongitude + decimalLatitude
rasValue = raster::extract(temp_raster, data_to_match) # layer = c(1:12)
combinePointValue = cbind(data_to_match,rasValue)

DT_1980s <- as.data.frame(combinePointValue)
# colnames(DT_1980s) <- gsub("^c\\..*", "Temp_1981_3", colnames(DT_1980s))
```

## aggregate monthly data
```{r}
DT_1980s %>% mutate(Temp1981 = rowMeans(select(DT_1980s, starts_with("X1981.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1982 = rowMeans(select(DT_1980s, starts_with("X1982.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1983 = rowMeans(select(DT_1980s, starts_with("X1983.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1984 = rowMeans(select(DT_1980s, starts_with("X1984.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1985 = rowMeans(select(DT_1980s, starts_with("X1985.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1986 = rowMeans(select(DT_1980s, starts_with("X1986.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1987 = rowMeans(select(DT_1980s, starts_with("X1987.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1988 = rowMeans(select(DT_1980s, starts_with("X1988.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1989 = rowMeans(select(DT_1980s, starts_with("X1989.")))) -> DT_1980s
DT_1980s %>% mutate(Temp1990 = rowMeans(select(DT_1980s, starts_with("X1990.")))) -> DT_1980s

DT_1980s %>% select(-starts_with("X")) -> DT_1980s

DT_1980s %>%
  mutate(Temp = case_when(
    year == 1981 ~ DT_1980s$Temp1981, 
    year == 1982 ~ DT_1980s$Temp1982, 
    year == 1983 ~ DT_1980s$Temp1983, 
    year == 1984 ~ DT_1980s$Temp1984, 
    year == 1985 ~ DT_1980s$Temp1985, 
    year == 1986 ~ DT_1980s$Temp1986, 
    year == 1987 ~ DT_1980s$Temp1987, 
    year == 1988 ~ DT_1980s$Temp1988, 
    year == 1989 ~ DT_1980s$Temp1989, 
    year == 1990 ~ DT_1980s$Temp1990)) -> DT_1980s                    

```

################################################################

# Extract data for 1990s
## Extract Temp values from RasterBrick
```{r}
data_to_match = D_1991_2000
temp_raster = temp1991_2000

coordinates(data_to_match)= ~ decimalLongitude + decimalLatitude
memory.limit(size=999999999)
rasValue = raster::extract(temp_raster, data_to_match) # layer = c(1:12)
combinePointValue = cbind(data_to_match,rasValue)

DT_1990s <- as.data.frame(combinePointValue)
# colnames(DT_1980s) <- gsub("^c\\..*", "Temp_1981_3", colnames(DT_1980s))
```

## aggregate monthly data
```{r}
DT_1990s %>% mutate(Temp1991 = rowMeans(select(DT_1990s, starts_with("X1991.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1992 = rowMeans(select(DT_1990s, starts_with("X1992.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1993 = rowMeans(select(DT_1990s, starts_with("X1993.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1994 = rowMeans(select(DT_1990s, starts_with("X1994.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1995 = rowMeans(select(DT_1990s, starts_with("X1995.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1996 = rowMeans(select(DT_1990s, starts_with("X1996.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1997 = rowMeans(select(DT_1990s, starts_with("X1997.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1998 = rowMeans(select(DT_1990s, starts_with("X1998.")))) -> DT_1990s
DT_1990s %>% mutate(Temp1999 = rowMeans(select(DT_1990s, starts_with("X1999.")))) -> DT_1990s
DT_1990s %>% mutate(Temp2000 = rowMeans(select(DT_1990s, starts_with("X2000.")))) -> DT_1990s

DT_1990s %>% select(-starts_with("X")) -> DT_1990s

DT_1990s %>%
  mutate(Temp = case_when(
    year == 1991 ~ DT_1990s$Temp1991, 
    year == 1992 ~ DT_1990s$Temp1992, 
    year == 1993 ~ DT_1990s$Temp1993, 
    year == 1994 ~ DT_1990s$Temp1994, 
    year == 1995 ~ DT_1990s$Temp1995, 
    year == 1996 ~ DT_1990s$Temp1996, 
    year == 1997 ~ DT_1990s$Temp1997, 
    year == 1998 ~ DT_1990s$Temp1998, 
    year == 1999 ~ DT_1990s$Temp1999, 
    year == 2000 ~ DT_1990s$Temp2000)) -> DT_1990s                    

```

################################################################

# Extract data for 2000s
## Extract Temp values from RasterBrick
```{r}
data_to_match = D_2001_2010
temp_raster = temp2001_2010

coordinates(data_to_match)= ~ decimalLongitude + decimalLatitude
rasValue = raster::extract(temp_raster, data_to_match) # layer = c(1:12)
combinePointValue = cbind(data_to_match,rasValue)

DT_2000s <- as.data.frame(combinePointValue)
```

## aggregate monthly data
```{r}
DT_2000s %>% mutate(Temp2001 = rowMeans(select(DT_2000s, starts_with("X2001.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2002 = rowMeans(select(DT_2000s, starts_with("X2002.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2003 = rowMeans(select(DT_2000s, starts_with("X2003.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2004 = rowMeans(select(DT_2000s, starts_with("X2004.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2005 = rowMeans(select(DT_2000s, starts_with("X2005.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2006 = rowMeans(select(DT_2000s, starts_with("X2006.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2007 = rowMeans(select(DT_2000s, starts_with("X2007.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2008 = rowMeans(select(DT_2000s, starts_with("X2008.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2009 = rowMeans(select(DT_2000s, starts_with("X2009.")))) -> DT_2000s
DT_2000s %>% mutate(Temp2010 = rowMeans(select(DT_2000s, starts_with("X2010.")))) -> DT_2000s

DT_2000s %>% select(-starts_with("X")) -> DT_2000s

DT_2000s %>%
  mutate(Temp = case_when(
    year == 2001 ~ DT_2000s$Temp2001, 
    year == 2002 ~ DT_2000s$Temp2002, 
    year == 2003 ~ DT_2000s$Temp2003, 
    year == 2004 ~ DT_2000s$Temp2004, 
    year == 2005 ~ DT_2000s$Temp2005, 
    year == 2006 ~ DT_2000s$Temp2006, 
    year == 2007 ~ DT_2000s$Temp2007, 
    year == 2008 ~ DT_2000s$Temp2008, 
    year == 2009 ~ DT_2000s$Temp2009, 
    year == 2010 ~ DT_2000s$Temp2010)) -> DT_2000s                    

```

################################################################

# Extract data for 2010s
## Extract Temp values from RasterBrick
```{r}
data_to_match = D_2011_2020
temp_raster = temp2011_2020

coordinates(data_to_match)= ~ decimalLongitude + decimalLatitude
rasValue = raster::extract(temp_raster, data_to_match) # layer = c(1:12)
combinePointValue = cbind(data_to_match,rasValue)

DT_2010s <- as.data.frame(combinePointValue)
```

## aggregate monthly data
```{r}
DT_2010s %>% mutate(Temp2011 = rowMeans(select(DT_2010s, starts_with("X2011.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2012 = rowMeans(select(DT_2010s, starts_with("X2012.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2013 = rowMeans(select(DT_2010s, starts_with("X2013.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2014 = rowMeans(select(DT_2010s, starts_with("X2014.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2015 = rowMeans(select(DT_2010s, starts_with("X2015.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2016 = rowMeans(select(DT_2010s, starts_with("X2016.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2017 = rowMeans(select(DT_2010s, starts_with("X2017.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2018 = rowMeans(select(DT_2010s, starts_with("X2018.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2019 = rowMeans(select(DT_2010s, starts_with("X2019.")))) -> DT_2010s
DT_2010s %>% mutate(Temp2020 = rowMeans(select(DT_2010s, starts_with("X2020.")))) -> DT_2010s

DT_2010s %>% select(-starts_with("X")) -> DT_2010s

DT_2010s %>%
  mutate(Temp = case_when(
    year == 2011 ~ DT_2010s$Temp2011, 
    year == 2012 ~ DT_2010s$Temp2012, 
    year == 2013 ~ DT_2010s$Temp2013, 
    year == 2014 ~ DT_2010s$Temp2014, 
    year == 2015 ~ DT_2010s$Temp2015, 
    year == 2016 ~ DT_2010s$Temp2016, 
    year == 2017 ~ DT_2010s$Temp2017, 
    year == 2018 ~ DT_2010s$Temp2018, 
    year == 2019 ~ DT_2010s$Temp2019, 
    year == 2020 ~ DT_2010s$Temp2020)) -> DT_2010s                    

```

################################################################################

# Combine the decade-dataframes
 
filter specific columns if needed - at least throw out the year-mean-temperature columns (col 23-32)
This is necessary since they differ between the decade-wise datasets
```{r}
# colnames(DT_1980s)
# DT_1980s %>% select(kingdom, order, species, id.grp, decimalLatitude, decimalLongitude, decade, year, doy, Temp) -> DT_1980s
# DT_1990s %>% select(kingdom, order, species, id.grp, decimalLatitude, decimalLongitude, decade, year, doy, Temp) -> DT_1990s
# DT_2000s %>% select(kingdom, order, species, id.grp, decimalLatitude, decimalLongitude, decade, year, doy, Temp) -> DT_2000s
# DT_2010s %>% select(kingdom, order, species, id.grp, decimalLatitude, decimalLongitude, decade, year, doy, Temp) -> DT_2010s

DT_1970s <- DT_1970s[,-c(23:32)]
DT_1980s <- DT_1980s[,-c(23:32)]
DT_1990s <- DT_1990s[,-c(23:32)]
DT_2000s <- DT_2000s[,-c(23:32)]
DT_2010s <- DT_2010s[,-c(23:32)]


DT <- rbind(DT_1970s, DT_1980s, DT_1990s, DT_2000s, DT_2010s)
```

################################################################################

# Save finishes dataframe object to a file

```{r}
saveRDS(DT, file = "DT2.rds")
```


# Rename columns
```{r}
names(DT)[names(DT)=="decimalLatitude"] <- "lat"
names(DT)[names(DT)=="decimalLongitude"] <- "long"
names(DT)[names(DT)=="Temp"] <- "temp"
```

# throw out all datapoints outside Germany


## limit (roughly) to Germany extensions 
```{r}
# DT %>% filter(lat >=47.2 & lat <= 54.9 & long >= 5.96 & long <= 15.02) -> DT
range(DT$lat)
range(DT$long)
```

## Load Germany shapefile
```{r}
library(rworldmap)
library(rgeos)
library(sf)

# Get the world map
worldMap <- getMap(resolution = "high")
GerMap <- worldMap[which(worldMap$NAME=="Germany"),]
plot(GerMap)
```

## map points
```{r}
GerMap_sf <- st_as_sf(GerMap)
ggplot(data = GerMap_sf)+
  geom_sf()+
  geom_point(data = DT, aes(y = lat, x = long))+
  coord_sf(xlim = c(0,16), ylim = c(45,55))
```
there are multiple points outside of Germany

## create a spatial points dataframe and cut all points outside of Germany 
```{r}
library("sp")
# define the spatial points (here the sampling points)
spatial_points <- cbind(DT$long, DT$lat)
spatial_points <- SpatialPoints(spatial_points, proj4string=CRS(proj4string(GerMap)))

# make a SpatialPointsDataFrame
DT_SP <- DT
coordinates(DT_SP)= ~ long + lat

# assign Coordinate reference system
CRS.new<-CRS("+init=epsg:4326")
proj4string(DT_SP) <- CRS.new 
proj4string(GerMap) <- CRS.new

# clip all points outside Germany
DT_SP <- DT_SP[GerMap,]
DT_test <- as.data.frame(DT_SP)


```

## Plot the refinded dataframe again 
now all points should be within the boarders of Germany
```{r}
GerMap_sf <- st_as_sf(GerMap)

library(ggplot2)
ggplot(data = GerMap_sf)+ 
  #geom_sf()+
  geom_point(data = DT_test, aes(y = lat, x = long))+
  coord_sf(xlim = c(0,16), ylim = c(47.24,55))
```


# Check N per decade/species ect AGAIN

sinse datapoints outside Germany where removed
check if there are still at least 10 datapoints per decade for each species
```{r}
N_per_decade2 <- DT_test %>% group_by(kingdom, order, species, decade2) %>% summarize(N = length(doy), mean = mean(doy))

# select species with at least 10 datapoints per decade
N_per_decade2 %>% filter(N > 9) -> N_per_decade2_min
# select species that have 4 decades of data, 2010s-decade including 2020
N_per_decade2_min %>% filter(n() > 3) -> N_per_decade2_min4 

dim(N_per_decade2_min[N_per_decade2_min$decade2=="1980",])
dim(N_per_decade2_min[N_per_decade2_min$decade2=="1990",])
dim(N_per_decade2_min[N_per_decade2_min$decade2=="2000",])
dim(N_per_decade2_min[N_per_decade2_min$decade2=="2010",])

length(unique(N_per_decade2_min4$species)) 
  
```
# Filter out all species that have now less than 10 datapoints per decade AGAIN
```{r}
Species_list2 <- unique(N_per_decade2_min4$species)
DT_test %>% filter(species %in% Species_list2) -> DF_test_sub
DF_test_sub <- droplevels(DF_test_sub)
length(levels(as.factor(DF_test_sub$species)))
```


## save the dataset
```{r}
saveRDS(DT_test, file = "D:/Franziska/CleanWorkBench/GBIF_Data/DT_after_nonGer_exclusion.rds")
saveRDS(DF_test_sub, file = "D:/Franziska/CleanWorkBench/GBIF_Data/DT_after_nonGer_exclusion_SpeciesFilteredAgain.rds") # this is the dataset where species were filtered again
```

```{r}
rm(CRS.new, DT_SP, DT_test, spatial_points, world, worldMap, DT_test_sub)
rm(df_sub, test, D,
   D_1971_1980, D_1981_1990, D_1991_2000, D_2001_2010, D_2011_2020, 
   data_to_match, combinePointValue, resValue, 
   DT, DT_1970s, DT_1980s, DT_1990s, DT_2000s, DT_2010s,
   temp_raster, temp1971_1980, temp1981_1990, temp1991_2000, temp2001_2010, temp2011_2020)

```


################################################################################

# Basic plots

```{r}
ggplot(data = DT_1980s, aes(year, Temp))+
  geom_point()+
  geom_smooth(method="lm")
```


```{r}
ggplot(data = DT_1980s, aes(Temp, doy, group = species))+
  geom_point()+
  geom_smooth(method="lm")
```


```{r}
ggplot(data = DT_1980s[DT_1980s$order=="Hymenoptera",], aes(Temp, doy, colour = species))+
  geom_point()+
  geom_smooth(method="lm", se = F)
```

